LaTeX style file for ESANN manuscripts

### 2017-11-29 ESSAN/article.tex
Modified diagrams and plot futher

\begin{figure}[h!]
\centering
\includegraphics[scale=1.5]{diagram.eps}
\caption{In this diagram we can appreciate the overall connectivity of the system. Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
\end{figure}

@@

In order to recall a particular sequence we cue the first element. This means that we artificially activate the first unit (clamping) of the sequence for a short time. After this initial cue, if the weights are set right, the neural network continues evolving on its own and a particular sequence is recalled. In figure \ref{Fig:recall} a successful recall process is shown.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.20]{recall.eps}
\caption{A succesful recall. a) The activity of the units as it evolves in time. b) The currents that each unit receives as the sequence progresses. c) the z-filters activations for each unit. d) The connectivity matrix $W$. The weights of the matrix can be divided on the excitatory connections: the ones that induce transitions $Exc_{T}$ and the ones that stabilize the unit $Exc_{self}$ and the inhibitory connections $Inh$.}\label{Fig:recall}
\end{figure}

@@

It is now instructive to detail at the level of the dynamics how the sequential activity is propagated. After the first unit $x_1$ has been activated for long enough due to clamping the filter $z_1$ will start catching up with the activity by increasing its value. When this happens the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$. As this process continues there will a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated. This, will again will prompt the filter $z_2$ to start catching up initializing the process described again. On the other hand as soon as the second filter starts growing $z_2$ the first unit will start receiving an inhibitory current $Inh \times z_2$ due to the backward inhibitory projection, this will in turn suppress the unit $x_1$ allowing the sequence to propagate. Finally, once the unit $x_1$ is finally suppressed the excitatory current mediated by $Exc_T$ will be diminished which in most cases would also suppress the activity of $x_2$ if were not for the self-sustaining current $Exc_{self} \times z_2$. In short, the term $Exc_T$ ensures the transition of the activity from one unit of the sequence to the next whereas the competition between $Exc_{self}$ and $Inh$ first lock the pattern unit and then finally suppress it allowing the sequence to evolve towards its end without hindrance.

With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics. First, if the magnitude of $Exc_{self}$ the activity will not be big enough to overcome the threshold and the sequence will not propagate. Second, if $Exc_{Self}$ is enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate. If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not enough to suppress the units the result will be that all the units in the sequence will be activated at the end.

@@

A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the required context. In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where the recall time is the amount of time that a particular unit is active during a sequence recall. In our network there are a two parameters that a priori are strong candidates to regulate the recall time: the transition weight $Exc_T$ by controlling how fast you activate the next unit in the sequence and the threshold $\theta$ by controlling how long it takes for the current to activate the next unit in a sequence. Here we show the exact results for $Exc_T$ but the reasoning for $\theta$ is analogous.

In order to calculate the recall time we make the following consideration. Once the unit $x_1$ becomes activated the filter $z_1$ starts increasing its value and the unit $x_2$ will be receiving an excitatory current $Exc_T \times z_1$. The point in time t at which the this current overcomes the threshold will be the recall time. Using that reasoning we can solve the following equation to calculate the recall time: $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$. If we do we end up with the following expression: $T_{recall} = \tau_z \log(\frac{Exc_T}{Exc_T - \theta})$. This expression tells us that we actually have a singularity at $Exc_T = \theta$, which for our purposes means that can make the recall time arbitrarily by setting this two parameters close enough. In our neural network however, there are more currents into play because of the all-to-all connectivity, we can include those corrections and we end up with an expression of the form $T_{recall} = \tau_z \log(\frac{Exc_T + Inh}{Exc_T - \theta})$ which also possess the after-mentioned property. We can see an illustration of the wide dynamical range of our system in figure \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.

@@

\begin{align}
\tag{pre-synaptic}
\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (1 - z_{post})z_{pre} \label{eq:pre-synaptic rule}  \\
\tag{post-synaptic}
\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
\end{align}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=1.3]{plasticity_diagram.eps}
\caption{A scheme of how plasticity works. a) Hebbian term contribution. b) The contribution is negative when the post-synaptic unit is active. c) The contribution is negative when the pre-synaptic unit is active. d) If units are not active there is no decay of the weights.}\label{Fig:plasticity diagram}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.3]{training_rule.eps}
\caption{Training with the learning rule. a) An example of training with a small value of $tau_z$. b) Same procedure as in a but with a longer time constant $\tau_z$. c) evolution of weights with the number of epochs. The three situations depicted utilize the pre-synaptic learning rule.}\label{Fig:epochs}
\end{figure}

>>>>

\begin{figure}[h!]
\centering
\includegraphics[scale=1.5]{diagram.eps}
\caption{Scheme of the overall network connectivity of the system. Note that the inhibitory connections are only fully depicted for the first unit for the sake of clarity.}\label{Fig:diagram}
\end{figure}

@@

In order to recall a particular sequence we cue the first element. This means that we artificially activate the first unit (clamping) of the sequence for a short time. After this initial cue, if the weights have the right configuration, the neural network will continue evolving on its own until the whole particular sequence is correctly recalled as shown in figure  \ref{Fig:recall}.

\begin{figure}[h!]
\centering
\includegraphics[scale=0.20]{recall.eps}
\caption{Demostration of the successful sequence recall with the corresponding connectivity matrix. a) The activity of the units over time. b) The currents that each unit receives as the sequence progresses. c) the z-filters activations for each unit. d) The connectivity matrix $W$. The weights of the matrix $W$ can be divided into the excitatory connections: the ones on the diagonal that stabilize the unit $Exc_{self}$, the ones under the diagonal that induce transitions $Exc_{T}$ and finally the ones that stabilize the unit  and the inhibitory connections $Inh$ which dominate the matrix $W$. e) The legend with the color codes, we use different colors and thickness to represent the different variables corresponding to different units. }\label{Fig:recall}
\end{figure}

@@

It is now instructive to detail at the level of the dynamics how the sequential activity is propagated. After the first unit $x_1$ has been activated for long enough due to clamping the filter $z_1$ will start catching up with the activity by increasing its value. It follow then that the unit corresponding to $x_2$ will receive a growing current with value $Exc_{T} \times z_1$ because of the forward connectivity. As this process continues there will be a point in time (exactly at $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$) when $x_2$ will receive enough current to become activated. This in turn will prompt the filter $z_2$ to start catching up initializing the describe process again. On the other hand as soon as the second filter starts growing $z_2$ the first unit will start receiving an inhibitory current $Inh \times z_2$ due to the backward inhibitory projection, this will in turn suppress the unit $x_1$ allowing the sequence to propagate. Finally, once the unit $x_1$ is finally suppressed the excitatory current mediated by $Exc_T$ will be diminished which in most cases would also suppress the activity of $x_2$ if were not for the self-sustaining current $Exc_{self} \times z_2$. In short, the term $Exc_T$ ensures the transition of the activity from one unit of the sequence to the next whereas the competition between $Exc_{self}$ and $Inh$ first lock the pattern unit and then finally suppress it allowing the sequence to evolve towards its end without hindrance.

With the knowledge above in mind we can envision the following  failure scenarios for the sequential dynamics. First, if the magnitude of $Exc_{self}$ is small the self-excitatory current will not be big enough to overcome the threshold and the sequence will not propagate. Second, if $Exc_{Self}$ is big enough to keep the unit active but $Exc_{T}$ is not big enough the next unit in the sequence will not be activated and the sequence will not propagate, in this case we are stuck in the first pattern forever. If both $Exc_{self}$ is big enough to sustain the activity and $Exc_{T}$ is big enough to propagate it but the inhibitory connection $Inh$ is not big enough to suppress the units the result will be that all the units in the sequence will be activated at the end.

@@

A common problem in sequence learning is to scale the time duration of the sequence according to the needs of the context. In terms of a neural network it means that once the dynamics of sequence recalling are stored in the shape of the weights we can control the \textit{recall time} of the sequence with the internal parameters of the network where we define the recall time as the amount of time that a particular unit is active during a sequence retrieval. In our network there are a two parameters that a priori are strong candidates to regulate the recall time: the transition weight $Exc_T$ by controlling how fast you activate the next unit in the sequence and the threshold $\theta$ by controlling how long it takes for the current to activate the next unit in a sequence. Here we show the exact results for $Exc_T$ but the reasoning for $\theta$ is analogous.

In order to calculate the recall time we make the following consideration. Once the unit $x_1$ becomes activated the filter $z_1$ starts increasing its value and the unit $x_2$ will be receiving an excitatory current $Exc_T \times z_1$. The point in time $t_{recall}$ at which the this current overcomes the threshold will be the recall time. Using that reasoning we can solve the following equation to calculate the recall time: $Exc_{T}(1 - e^{-\frac{t}{\tau_z}}) - \theta = 0$. If we do we end up with the following expression: $T_{recall} = \tau_z \log(\frac{Exc_T}{Exc_T - \theta})$. This expression tells us that we actually have a singularity at $Exc_T = \theta$, which for our purposes means that can make the recall time arbitrarily by setting this two parameters close enough. In our neural network however, there are more currents into play because of the all-to-all connectivity, we can include those corrections and we end up with an expression of the form $T_{recall} = \tau_z \log(\frac{Exc_T + Inh}{Exc_T - \theta})$ which also possess the after-mentioned property. We can see an illustration of the wide dynamical range of our system in figure \ref{Fig:dynamical_range} where we also show two examples of sequence recall with different $Exc_T$ values and therefore with different recall times.

@@

\begin{align}
\tag{pre-synaptic}
\tau_w \dfrac{dw}{dt} &= (w_{max} - w) z_{pre}z_{post} + (w_{min} - w) (1 - z_{post})z_{pre} \label{eq:pre-synaptic rule}  \\
\tag{post-synaptic}
\tau_w \dfrac{dw}{dt} &= \underbrace{(w_{max} - w)}_{\text{saturation}} \underbrace{z_{pre}z_{post}}_{\text{Hebbian}} + \underbrace{(w_{min} - w)}_{\text{saturation}}\underbrace{ (1 - z_{pre})z_{post}}_{\text{synaptic-gating}} \label{eq:post-synaptic rule}
\end{align}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=1.3]{plasticity_diagram.eps}
\caption{Conceptual illustration of the mechanism of learning induced synaptic changes. Here the a black filling stands for an active unit a) Hebbian associative plasticity that implies a weight increase when both pre- and pos-synaptic units are simultaneously active b) In the case of the post-synaptic rule the weight is decreased if the post-synaptic unit itself active without the pre-synaptic unit being active. c) In the case of the pre-synaptic rule the weight is decreased if the pre-synaptic unit is itself active without the post-synaptic unit being active. d) If neither pre- nor post-synaptic units are active, there is no weight modification.}\label{Fig:plasticity diagram}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.3]{training_rule.eps}
\caption{The effect of training with the learning rule. a) An example of training with a small value of $tau_z=0.050s$. b) Same procedure as in a but with a longer time constant $\tau_z=0.150s$. c) evolution in time of the connection weights from the first to the second unit $Exc_T$, the one from the second back to the first unit $Inh$ and finally of the second unit into itself $Exc_{self}$. Five epochs of length $2s$ are shown here, note that this is enough time for the weights to stabilize.}\label{Fig:epochs}
\end{figure}
